The all-seeing eye of resonant Auger electron spectroscopy: a study on aqueous KCl

### 2018-04-23 abstract.tex
apr 23
preparation of jpcl draft started

X-ray absorption and Auger electron spectroscopies are powerful tools to probe the electronic structure and immediate surroundings of ions in a solution. In this work we use a combination of these methods to study the electronic structure and decay in aqueous KCl at the K-edges of \ki~and \cli. Although the two ions are isoelectronic, their Auger electron spectra as a function of the photon energy exhibit notably different features. To explain these differences, we carried out {\it ab initio} calculations of both the core excited states and the final Auger states of bare \ki, \cli~and their microsolvated clusters. Our calculations show that the energy order of the 3d and 4p orbitals is inverted in \ki~with respect to \cli. The reverse orbital order in the two ions is reflected in both the ordering of the core excited states and in the final states populated in the resonant Auger decay. The energetic proximity of the 3d and 4p virtual states in the bare \ki~ion leads to their mixing in the presence of the solvent, and to the population of the dipole forbidden 1s$^{-1}$3d state upon K-shell excitation in an aqueous solution. The resonant Auger decay of this state results in a separate feature in the Auger electron spectrum of \ki~which is absent in the spectrum of \cli. The results of this work represent a pioneering study of the decay processes initiated by photoabsorption in the tender x-ray regime close to threshold in liquids and are thus of importance in unveiling the mechanisms of radiation damage in biologically relevant systems.
%Denis: line 2, I replaced spectroscopies by methods

>>>>

X-ray absorption and Auger electron spectroscopies are powerful tools to probe the electronic structure and immediate surroundings of ions in solution. In this work we use a combination of these methods together with {\it ab initio} calculations to study the electronic structure and decay in aqueous KCl at the K-edges of \ki~and \cli. Although the two ions are isoelectronic, their Auger electron spectra exhibit notably different features. %To explain these differences, we carried out {\it ab initio} calculations of both the core excited states and the final Auger states of bare \ki, \cli~and their microsolvated clusters. 
This is due to the fact that the energetic order of the 1s$^{-1}$3d and 1s$^{-1}$4p core excited states is inverted in \ki~with respect to \cli, and moreover the two states lie energetically close in the bare \ki~ion leading to their mixing in the presence of the solvent. As a result, the dipole forbidden 1s$^{-1}$3d state is populated upon K-shell excitation of aqueous K$^{+}$. The resonant Auger decay of this state results in a separate feature in the Auger electron spectrum of \ki~which is absent in the spectrum of \cli. The results of this work represent a pioneering study of the decay processes initiated by photoabsorption in the tender x-ray regime close to threshold in liquids and are thus of importance in unveiling the mechanisms of radiation damage in biologically relevant systems.
%Denis: line 2, I replaced spectroscopies by methods

### 2018-04-23 conclusion.tex
apr 23
preparation of jpcl draft started

\section{Conclusion}\label{sec:concl}

Using a combination of x-ray absorption and Auger electron spectroscopy in the tender x-ray regime, we studied the electronic structure of aqueous solution of KCl at the K-edges of both K and Cl. The Auger electron spectra of both ions as a function of photon energy exhibit features of normal as well as resonant Auger processes. To interpret the resonant Auger features in the experimental spectrum, we performed {\it ab initio} calculations on microsolvated clusters of \ki~and \cli. Our calculations show that the energy ordering of the 3d and 4p virtual orbitals of \cli~is inverted compared to \ki, and also that the energy splitting between the bright 1s$\,\rightarrow\,$4p and dark 1s$\,\rightarrow\,$3d core excited states is larger in the chloride case. Thus, the energetic proximity of the 3d and 4p orbitals in the bare \ki~ion results in the dipole forbidden 1s$\,\rightarrow\,$3d state acquiring intensity in a solution as a result of mixing with the dipole allowed 1s$\,\rightarrow\,$4p excitation. The spectator Auger decay of this state produces an additional dispersive feature which is manifested as a separate peak in the Auger electron spectrum at high kinetic energies. In the case of \cli~the 4p and 3d core excited states do not interact, and therefore, only fingerprints of the population and Auger decay of the dipole allowed 1s$\,\rightarrow\,$4p state are observed in the spectrum. Moreover, using the core-hole clock method we estimated the time of delocalization of the core excited electron at the pre-edge region of \cli$_{\text{aq}}$. Our results show that in this case, the resonant Auger decay and the delocalization of the excited electron occur on a comparable timescale. In the case of \ki$_{\text{aq}}$, we cannot make an accurate estimate, however, one can expect a much less efficient delocalization of the core excited electron.

Our work shows that the combination of x-ray absorption and resonant Auger spectroscopies is a sensitive probe of the electronic structure of solvated ions. The reported results are an important first step in the study of the electronic decay processes following photoabsorption in the tender x-ray regime, and they can have implications in revealing the mechanisms of radiation damage in biologically relevant systems.

>>>>

%\section{Conclusion}\label{sec:concl}

In summary, we studied the electronic structure of aqueous solution of KCl at the K-edges of both K and Cl using a combination of x-ray absorption and Auger electron spectroscopy in the tender x-ray regime, and {\it ab initio} calculations. The Auger electron spectra of both ions as a function of photon energy exhibit features of normal as well as resonant Auger processes. The latter process proceeds differently for aqueous K$^{+}$ and Cl$^{-}$ due to the population of the dipole forbidden K$^{+}$ 1s$^{-1}$3d state in a solution. The spectator Auger decay of this state produces an additional dispersive feature which is manifested as a separate peak in the Auger electron spectrum. In the case of \cli~only fingerprints of the population and Auger decay of the dipole allowed 1s$^{-1}$4p excitation are observed in the spectrum.

%Using a combination of x-ray absorption and Auger electron spectroscopy in the tender x-ray regime, we studied the electronic structure of aqueous solution of KCl at the K-edges of both K and Cl. The Auger electron spectra of both ions as a function of photon energy exhibit features of normal as well as resonant Auger processes. To interpret the resonant Auger features in the experimental spectrum, we performed {\it ab initio} calculations on microsolvated clusters of \ki~and \cli. Our calculations show that the energy ordering of the 3d and 4p virtual orbitals of \cli~is inverted compared to \ki, and also that the energy splitting between the bright 1s$^{-1}$4p and dark 1s$^{-1}$3d core excited states is larger in the chloride case. Thus, the energetic proximity of the 3d and 4p orbitals in the bare \ki~ion results in the dipole forbidden 1s$^{-1}$3d state acquiring intensity in a solution as a result of mixing with the dipole allowed 1s$^{-1}$4p excitation. The spectator Auger decay of this state produces an additional dispersive feature which is manifested as a separate peak in the Auger electron spectrum at high kinetic energies. In the case of \cli~the 1s$^{-1}$4p and 1s$^{-1}$3d core excited states do not interact, and therefore, only fingerprints of the population and Auger decay of the dipole allowed 1s$^{-1}$4p excitation are observed in the spectrum. Moreover, using the core-hole clock method we estimated the time of delocalization of the core excited electron at the pre-edge region of \cli$_{\text{aq}}$. Our results show that in this case, the resonant Auger decay and the delocalization of the excited electron occur on a comparable timescale. In the case of \ki$_{\text{aq}}$, an accurate estimate is difficult, however, one can expect a much less efficient delocalization of the core excited electron.

Our work shows that the combination of x-ray absorption and resonant Auger spectroscopy is a sensitive probe of the electronic structure of solvated ions. The reported results are an important first step in the study of the chains of relaxation steps triggered by photoabsorption in the tender x-ray regime. The Auger processes considered here are inevitably followed by multiple intra- and interatomic electronic decays, such as interatomic Coulombic decay (ICD) and electron-transfer mediated decay (ETMD). As a result of the latter processes, genotoxic free radicals and slow electrons are formed in the vicinity of the metal center. The magnitude of the damage inflicted upon the environment and the energies of the emitted electrons depend on the initial Auger step, and can therefore be controlled by tuning the energy of the radiation. Consequently, the results of this work can have implications in understanding radiation chemistry and radiation damage in biologically relevant systems.

### 2018-04-23 introduction.tex
apr 23
preparation of jpcl draft started

\section{Introduction}

X-ray absorption and Auger electron spectroscopy are powerful tools to study the electronic structure and the nearest environment of atoms and molecules in gas, liquid and solid phase. Understanding how atoms or molecules respond to irradiation with x-rays gives insight into the structure of solutions (see Ref.\ \citep{smith17:13909} and references therein), and the mechanism of radiation damage \citep{ONeill02:329,Carugo05:213,Stumpf16:237}. Upon absorption of an x-ray photon, either core excited or core ionized states of a specific atom are populated depending on the photon energy. The relaxation of these highly energetic states involves an ultrafast cascade of intraatomic processes, such as radiative and Auger decays. Furthermore, if the initially excited or ionized species is embedded in an environment, interatomic processes such as charge and energy transfer \citep{Pokapanich09:7264,Pokapanich11:13430,Stumpf16:237,unger17:708,ceolin17:263003} are possible.

The course of a decay cascade depends on the character of the initially populated states. This has been well understood in atoms and molecules in gas phase  \citep{stoychev08:074307,Demekhin08:043421,Demekhin09:104303,Ouchi11:053415,Miteva14:164303,travnikova16:213001,Gokhberg14:661,Trinter14:664}. In the case of a core ionized state, the Auger decay process, designated as normal Auger decay (see Fig.\ \ref{fg:auger}), leads to the population of doubly ionized final states localized on the initially ionized unit \citep{stoychev08:074307,Demekhin08:043421,Demekhin09:104303,Ouchi11:053415}.  The normal Auger decay process in clusters proceeds similarly to that in atoms or molecules. However, in the case of a core excited state, the resonant Auger process competes with the process of delocalization of the excited electron in clusters. If the initially core excited electron delocalizes within the lifetime of the core hole, then normal instead of resonant Auger decay is observed \citep{Bjorneholm95:3017}.

In a solution, the electronic decay processes initiated by x-ray photoabsorption are different compared to those in rare gas clusters due to the shorter distances and stronger interatomic interactions. In particular, the solvent molecules have two effects -- first, they strongly affect the excited \citep{miteva16:16671} or ionized states of the ion and second, they can participate in the decay processes, leading to the population of charge-separated final states, and ionization of the surrounding environment \citep{Pokapanich09:7264,Pokapanich11:13430,Stumpf16:237,ceolin17:263003,unger17:708}. Moreover, the process of delocalization of the initially excited electron also occurs in aqueous solutions \citep{Nordlund07:217406,Ottosson11:13489}. In the case of pure water, the delocalization of the O1s excited electron takes place on a femtosecond to sub-femtosecond time scale depending on the photon energy, thus being commensurate with the lifetime of the O1s core hole which is $\sim$3.6\,fs \citep{Nordlund07:217406}. The O K-edge is located in the soft x-ray range of photon energies, $\sim$533\,eV in water vapor \citep{wagner80:1445}. Going higher in photon energy, in the tender and hard x-ray regimes, the lifetimes of the core ionized or core excited states become even shorter, on the order of 1\,fs. And thus, it is even more imperative to reveal whether the delocalization of the core excited electron occurs within the lifetime of the core hole.

The aim of this work is to elucidate the nature of the states populated upon x-ray irradiation of solvated ions in the tender x-ray regime, and furthermore, to understand whether the process of delocalization influences the resonant Auger decay. To this end, we used Auger electron spectroscopy together with x-ray absorption spectroscopy in the tender x-ray regime to study aqueous potassium chloride at the K-edges of both \ki~and \cli. In particular, we demonstrate experimentally that at photon energies below the K-edges of the two ions, localized core excited states are populated. These states undergo resonant Auger decay within less than 1\,fs. In both ions, there is a competition between resonant Auger decay and delocalization of the excited electron. Using the core-hole clock method we show that in \ki~delocalization at the pre-edge is weak, whereas in the case of \cli, due to the energetic proximity of the core excited state to the ionization threshold, the rate of delocalization is of the same order as that of the resonant Auger process. Moreover, we observe that although the \ki~and \cli~ions are isoelectronic, they have different fingerprints in the resonant Auger spectra. With the aid of high-level {\it ab initio} calculations of the initial and final states of the resonant Auger process of both the bare ions and their microsolvated clusters, we demonstrate that these differences result from different electronic structures of the two ions, thus confirming that the combination of XAS and AES techniques is a sensitive probe of the electronic structure of solutions.

>>>>

%\section{Introduction}

X-ray absorption (XAS) and Auger electron spectroscopies (AES) are powerful tools to study the electronic structure and the nearest environment of atoms and molecules in gas, liquid and solid phase. Understanding how atoms or molecules respond to irradiation with x-rays gives insight into the structure of solutions (see Ref.\ \citep{smith17:13909} and references therein), and the mechanism of radiation damage \citep{ONeill02:329,Carugo05:213,Stumpf16:237}. Upon absorption of an x-ray photon, either core excited or core ionized states of a specific atom are populated depending on the photon energy. The relaxation of these highly energetic states involves an ultrafast cascade of intraatomic processes, such as radiative and Auger decays. Furthermore, if the initially excited or ionized species is embedded in an environment, interatomic processes such as charge and energy transfer \citep{Pokapanich09:7264,Pokapanich11:13430,Stumpf16:237,unger17:708,ceolin17:263003} are possible.

It has been well understood in atoms and molecules in gas phase that the course of a decay cascade depends on the character of the initially populated states \citep{stoychev08:074307,Demekhin08:043421,Demekhin09:104303,Ouchi11:053415,Miteva14:164303,travnikova16:213001,Gokhberg14:661,Trinter14:664}.
%The course of a decay cascade depends on the character of the initially populated states. This has been well understood in atoms and molecules in gas phase  \citep{stoychev08:074307,Demekhin08:043421,Demekhin09:104303,Ouchi11:053415,Miteva14:164303,travnikova16:213001,Gokhberg14:661,Trinter14:664}. In the case of a core ionized state, the Auger decay process, designated as normal Auger decay (see Fig.\ \ref{fg:auger}), leads to the population of doubly ionized final states localized on the initially ionized unit \citep{stoychev08:074307,Demekhin08:043421,Demekhin09:104303,Ouchi11:053415}.  The normal Auger decay process in clusters proceeds similarly to that in atoms or molecules. However, in the case of a core excited state, the resonant Auger process competes with the process of delocalization of the excited electron within the cluster. If the initially core excited electron delocalizes within the lifetime of the core hole, then normal instead of resonant Auger decay is observed \citep{Bjorneholm95:3017}. 
%
%
In a solution, the electronic decay processes initiated by x-ray photoabsorption are different compared to those in rare gas clusters due to the shorter distances and stronger interatomic interactions. In particular, the solvent molecules have two effects -- first, they strongly affect the excited \citep{miteva16:16671} or ionized states of the ion and second, they can participate in the decay processes, leading to the population of charge-separated final states, and ionization of the surrounding environment \citep{Pokapanich09:7264,Pokapanich11:13430,Stumpf16:237,ceolin17:263003,unger17:708}. Moreover, the process of delocalization of the initially excited electron also occurs in aqueous solutions \citep{Nordlund07:217406,Ottosson11:13489}. In the case of pure water, the delocalization of the O1s excited electron takes place on a femtosecond to sub-femtosecond time scale depending on the photon energy, thus being commensurate with the lifetime of the O1s core hole which is $\sim$3.6\,fs \citep{Nordlund07:217406}. The O K-edge is located in the soft x-ray range of photon energies, $\sim$533\,eV in water vapor \citep{wagner80:1445}. Going higher in photon energy, to the tender and hard x-ray regimes, the lifetimes of the core ionized or core excited states become even shorter, on the order of 1\,fs. And thus, it is even more imperative to reveal whether the delocalization of the core excited electron occurs within the lifetime of the core hole.

The aim of this work is to elucidate the nature of the states populated upon x-ray irradiation of solvated ions in the tender x-ray regime, and furthermore, to understand whether the process of delocalization influences the resonant Auger decay. To this end, we used Auger electron spectroscopy together with x-ray absorption spectroscopy in the tender x-ray regime to study aqueous potassium chloride at the K-edges of both \ki~and \cli. In particular, we demonstrate experimentally that at photon energies below the K-edges of the two ions, core excited states are populated. These states undergo resonant Auger decay within less than 1\,fs. In both ions, there is a competition between resonant Auger decay and delocalization of the excited electron. Using the core-hole clock method we show that in \ki~delocalization at the pre-edge is slower than the Auger process. However, in the case of \cli, the rate of delocalization is of the same order as that of the resonant Auger process due to the energetic proximity of the core excited state to the ionization threshold. Moreover, we observe that although the \ki~and \cli~ions are isoelectronic, they have different fingerprints in the resonant Auger spectra. With the aid of high-level {\it ab initio} calculations of the initial and final states of the resonant Auger process of both the bare ions and their microsolvated clusters, we demonstrate that these differences result from different electronic structures of the two ions, thus confirming that the combination of XAS and AES techniques is a sensitive probe of the electronic structure of solutions.

### 2018-04-23 master.tex
apr 23
preparation of jpcl draft started

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jacsat,manuscript=article]{achemso}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\title{The all-seeing eye of resonant Auger electron spectroscopy: a study on aqueous KCl}
\title{Tracking the hazy boundary between resonant and normal Auger decay near the K-edges of aqueous \ki~and \cli}

@@

We thank Prof.\ Nobuhiro Kosugi and Dr.\ Matja\v{z} \v{Z}itnik for the fruitful discussions. Experiments were performed at the GALAXIES beamline, SOLEIL Synchrotron, France (Proposal No. 20140160). The authors are grateful to the SOLEIL staff for assistance during the beamtime. This project has received funding from the Research Executive Agency (REA) under the European Union$'$s Horizon 2020 research and innovation programme Grant agreement No 705515. Campus France and the PHC SIAM exchange program are acknowledged for financial support (project No. 38282QB). N. V. Kryzhevoi and L. S. Cederbaum acknowledge the financial support of the European Research Council (ERC) (Advanced Investigator Grant No. 692657) and the Deutsche Forschungsgemeinschaft (DFG research unit 1789).

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jpclcd,manuscript=letter]{achemso}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{The all-seeing eye of resonant Auger electron spectroscopy: a study on aqueous KCl}
%\title{Tracking the hazy boundary between resonant and normal Auger decay near the K-edges of aqueous \ki~and \cli}

@@

We thank Prof.\ Nobuhiro Kosugi and Dr.\ Matja\v{z} \v{Z}itnik for the fruitful discussions. Experiments were performed at the GALAXIES beamline, SOLEIL Synchrotron, France (Proposal No.\ 20140160). The authors are grateful to the SOLEIL staff for assistance during the beamtime. This project has received funding from the Research Executive Agency (REA) under the European Union$'$s Horizon 2020 research and innovation programme Grant agreement No.\ 705515. Campus France and the PHC SIAM exchange program are acknowledged for financial support (project No.\ 38282QB). L. S. Cederbaum and N. V. Kryzhevoi acknowledge the financial support of the European Research Council (ERC) (Advanced Investigator Grant No.\ 692657) and the Deutsche Forschungsgemeinschaft (DFG research unit 1789).

### 2018-04-23 methods.tex
apr 23
preparation of jpcl draft started

\section{Methods} \label{sec:methods}
\subsection{Experimental}

For the present experiment we used the newly operational microjet setup that was specifically designed for the HAXPES station of the GALAXIES beamline \citep{ceolin13:188,rueff15:175}. A differentially-pumped tube in which the microjet head is inserted, is mounted on a 3-axes motorized manipulator in front of the spectrometer lens. Two holes of 2\,mm diameter allow the photons to go in and out. At the end of the tube and in front of the lens, a 500\,$\mu$m diameter hole skimmer allows the electrons created at the interaction point to go in the direction of the spectrometer. The microjet head is mostly composed of a 30\,$\mu$m diameter vertical glass capillary facing a temperature-controlled catcher in CuBe having a 300\,$\mu$m hole, and a camera. Piezo motors allow their precise alignment relative to each other and to the photon beam. The catcher is placed at a distance of about 5\,mm from the capillary and is permanently pumped in order to extract the liquid. For the present experiment, a 0.5M KCl aqueous solution is injected in the capillary by a high performance liquid chromatography (HPLC) pump with a constant flux of 1.6\,ml/min. The alignment of the setup is performed on the KCl aqueous solution by measuring the water O1s XPS peak intensity and by optimizing the liquid vs gas phase ratio. The pressure in the main chamber is kept below the 10$^{-5}$\,mbar range whereas it is kept at about 10$^{-4}$\,mbar in the differentially-pumped tube when the HPLC pump is ON. Our equipment is an updated version of the equipment used in Ref.\ \cite{faubel88:269}. The aqueous potassium chloride solution was prepared by mixing >99\% KCl salt with deionized water. Filtering and degazing procedures were systematically performed before injecting the solution. The spectrometer resolution of about 0.6\,eV was achieved with the 500\,eV pass energy and 0.5\,mm slits. The photon energy resolution achieved at 2.8\,keV and 3.6\,keV was about 0.3\,eV and 0.4\,eV, respectively. The experimental 2D maps representing the evolution of the KLL Auger spectra in the vicinity of the \cli and \ki~K-edges, as a function of the photon energy, are shown in Figs.\ \ref{fg:2dmap_k} and \ref{fg:2dmap_cl}, respectively. The aqueous \ki~and \cli~1s ionization potentials were measured at h$\nu$=5keV and calibrated on the liquid contribution of the O1s XPS spectrum \cite{winter06:1176}. The maps were also calibrated using the O1s photoelectron line of liquid water but at photon energies close to the potassium and chloride 1s ionization thresholds.

\subsection{{\bf{\it Ab initio}} calculations}

The theoretical X-ray absorption spectra were computed for the hexa-coordinated clusters of both ions, \ki(H$_2$O)$_6$ and \cli(H$_2$O)$_6$, which can be considered as representatives of the complete first solvation shell of the two ions \citep{Ohtaki93:1157,soper06:180,ma14:1006}. The two structures shown in Fig.\ \ref{fg:xas_kcl} were optimized at the DFT level of theory using the B3LYP functional and the 6-311++G(2d,2p) basis set \citep{Krishnan80:650,Blaudeau97:5016}. The geometry optimization was performed with the Gaussian 09 package \citep{g09}. In order to obtain realistic structures corresponding to the bulk solution, we carried out constrained geometry optimization by choosing the equilibrium gas-phase geometries \citep{lee99:3995,lee02:5509} belonging to the D$_3$ point group and then increasing the angle $\theta$ between the K-O/Cl-O bond and the $C_3$ axis to 55$^{\circ}$ in the case of \ki~and 50$^{\circ}$ in the case of \cli. This angle was chosen such that the O-K-O and O-Cl-O angles are around the maxima of the angular distributions obtained from quantum mechanics / molecular mechanics dynamical simulations in Ref.\ \citep{ma14:1006}. Moreover, we fixed the K-O and Cl-O distances to 2.840~\AA~and 3.140~\AA, respectively, such that they correspond to the distances from other theoretical and experimental works \citep{ge13:13169,gora00:7,Ohtaki93:1157,soper06:180,ma14:1006}.

The energies and transition moments of the core excited states of the bare ions and microsolvated clusters were computed with the algebraic diagrammatic construction method for the polarization propagator \citep{sch82:2395} within the core-valence separation approximation \citep{bar85:867,ced80:206,ced81:1038} (CVS-ADC(2)x) as implemented in the Q-Chem package \citep{Wenzel14:1900,Wenzel14:4583,Wormit14:774,QChem2015}. In the case of \cli~the 6-311++G(3df,3pd) basis set \citep{Krishnan80:650,McLean80:5639} (excluding f functions) was used for all atoms, whereas in the case of \ki~we used the 6-311+G(2d,p) basis set \citep{Krishnan80:650,Blaudeau97:5016} on all atoms, and two additional sets of s, p and d diffuse functions were added on K. The use of a smaller basis set in the case of K is due to the higher number of atomic orbitals compared to the case of Cl, and therefore, prohibitively high cost of the CVS-ADC(2)x computation. In our calculations the core space comprises the 1s orbital of K$^{+}$ or \cli, whereas the remaining occupied orbitals are included in the valence space. For the calculations of the XAS spectra we used the C$_2$ point group in the case of K$^{+}$(H$_2$O)$_6$ and \cli(H$_2$O)$_6$. To account for the lifetime broadening due to the Auger decay of the core excited states, we convolved the theoretical spectra with a Lorentzian function of FWHM 0.74\,eV and 0.62\,eV in the case of \ki~and \cli, respectively \citep{Krause79:329}. Additionally, we convolved the theoretical spectra with a Voigt profile to also account for the experimental resolution (see Fig.\ \ref{fg:xas_kcl}). We analyzed the core excited states by expanding the natural orbitals occupied by the excited electron (singly occupied natural orbitals, SONOs) $\psi_{i}$ of the microsolvated clusters in the basis of SONOs of the bare K$^{+}$ or \cli~ion $\chi_{nl}$
%
\begin{equation}\label{eq:sono_proj}
\psi_{i} = \sum_{nl} a^{i}_{nl} \chi_{nl}
\end{equation}
%
where $n$ and $l$ stand for the principal and orbital quantum numbers as described in Ref.\ \citep{miteva16:16671}. The expansion coefficients $a^{i}_{nl}$ show the degree of delocalization of the excited electron and the mixing of the core excited states in the ligand field created by the surrounding water molecules (see Fig.\ \ref{fg:xas_kcl}).

The final states following KLL resonant Auger decay of K$^{+}$(H$_2$O)$_6$ and \cli(H$_2$O)$_6$ were computed at the Configuration Interaction Singles (CIS) level using the Graphical Unitary Group Approach (GUGA) as implemented in the GAMESS-US package \citep{GUGA_PhysScr_21,GUGA_JCP_70,GUS}. In order to account for the relaxation effects upon core ionization, we used a restricted open-shell Hartree-Fock reference wave function with a hole in the 2s orbital of both \ki~and \cli.  We used the 6-311++G(2d,2p) basis set \citep{Krishnan80:650,McLean80:5639,Blaudeau97:5016} on all atoms. Additionally, the basis set was augmented with two sets of s, p, d diffuse functions in the case of \ki, and three sets of s, p, d diffuse functions in the case of \cli. The larger basis set employed in the case of Cl was necessary in order to ensure the convergence of the excited states. The active space comprises the 2s and 2p orbitals of K/Cl with occupancy fixed to 6 and all virtual orbitals with occupancy fixed to 1. The remaining doubly occupied orbitals were frozen in the calculation. \citep{mosnier16:061401}

>>>>

%\section{Methods} \label{sec:methods}
%\subsection{Experimental}

For the present experiment we used the newly operational microjet setup that was specifically designed for the HAXPES station of the GALAXIES beamline \citep{ceolin13:188,rueff15:175}. A differentially-pumped tube in which the microjet head is inserted, is mounted on a 3-axes motorized manipulator in front of the spectrometer lens. Two holes of 2\,mm diameter allow the photons to go in and out. At the end of the tube and in front of the lens, a 500\,$\mu$m diameter hole skimmer allows the electrons created at the interaction point to go in the direction of the spectrometer. The microjet head is mostly composed of a 30\,$\mu$m diameter vertical glass capillary facing a temperature-controlled catcher in CuBe having a 300\,$\mu$m hole, and a camera. Piezo motors allow their precise alignment relative to each other and to the photon beam. The catcher is placed at a distance of about 5\,mm from the capillary and is permanently pumped in order to extract the liquid. For the present experiment, a 0.5M KCl aqueous solution is injected in the capillary by a high performance liquid chromatography (HPLC) pump with a constant flux of 1.6\,ml/min. The alignment of the setup is performed on the KCl aqueous solution by measuring the water O1s x-ray photoelectron peak intensity and by optimizing the liquid vs gas phase ratio. The pressure in the main chamber is kept below the 10$^{-5}$\,mbar range whereas it is kept at about 10$^{-4}$\,mbar in the differentially-pumped tube when the HPLC pump is ON. Our equipment is an updated version of the equipment used in Ref.\ \cite{faubel88:269}. The aqueous potassium chloride solution was prepared by mixing >99\% KCl salt with deionized water. Filtering and degazing procedures were systematically performed before injecting the solution. The spectrometer resolution of about 0.6\,eV was achieved with the 500\,eV pass energy and 0.5\,mm slits. The photon energy resolution achieved at 2.8\,keV and 3.6\,keV was about 0.3\,eV and 0.4\,eV, respectively. The experimental 2D maps representing the evolution of the KLL Auger spectra in the vicinity of the \cli and \ki~K-edges, as a function of the photon energy, are shown in Figs.\ \ref{fg:2dmap_k} and \ref{fg:2dmap_cl}, respectively. The aqueous \ki~and \cli~1s ionization potentials were measured at h$\nu = 5$\,keV and calibrated on the liquid contribution of the O1s XPS spectrum \cite{winter06:1176}. The maps were also calibrated using the O1s photoelectron line of liquid water but at photon energies close to the potassium and chloride 1s ionization thresholds.

%\subsection{{\bf{\it Ab initio}} calculations}

The theoretical X-ray absorption spectra were computed for the hexa-coordinated clusters of both ions, \ki(H$_2$O)$_6$ and \cli(H$_2$O)$_6$, which can be considered as representatives of the complete first solvation shell of the two ions \citep{Ohtaki93:1157,soper06:180,ma14:1006}. The two structures shown in Fig.\ \ref{fg:xas_kcl} were optimized at the DFT level of theory using the B3LYP functional and the 6-311++G(2d,2p) basis set \citep{Krishnan80:650,Blaudeau97:5016}. The geometry optimization was performed with the Gaussian 09 package \citep{g09}. In order to obtain a realistic structure for \ki~corresponding to the bulk solution, we carried out constrained geometry optimization by choosing the equilibrium gas-phase geometries \citep{lee99:3995,lee02:5509} belonging to the D$_3$ point group and then increasing the angle $\theta$ between the K-O bond and the $C_3$ axis to 55$^{\circ}$. This angle was chosen such that the O-K-O angles are around the maxima of the angular distributions obtained from quantum mechanics/molecular mechanics simulations in Ref.\ \citep{ma14:1006}. Moreover, we fixed the K-O distance to 2.840~\AA, such that it corresponds to the distances from other theoretical and experimental works \citep{Ohtaki93:1157,soper06:180,ma14:1006}.
%\citep{ge13:13169,gora00:7,Ohtaki93:1157,soper06:180,ma14:1006}.
%Moreover, we fixed the K-O and Cl-O distances to 2.840~\AA~and 3.140~\AA, respectively, such that they correspond to the distances from other theoretical and experimental works \citep{ge13:13169,gora00:7,Ohtaki93:1157,soper06:180,ma14:1006}.

The energies and transition moments of the core excited states of the bare ions and microsolvated clusters were computed with the algebraic diagrammatic construction method for the polarization propagator \citep{sch82:2395} within the core-valence separation approximation \citep{bar85:867,ced80:206,ced81:1038} (CVS-ADC(2)x) as implemented in the Q-Chem package \citep{Wenzel14:1900,Wenzel14:4583,Wormit14:774,QChem2015}. In the case of \cli~the 6-311++G(3df,3pd) basis set \citep{Krishnan80:650,McLean80:5639} (excluding f functions) was used on all atoms, whereas in the case of \ki~we used the 6-311+G(2d,p) basis set \citep{Krishnan80:650,Blaudeau97:5016} on all atoms, and two additional sets of s, p and d diffuse functions were added on K. The use of a smaller basis set in the case of K is due to the higher number of atomic orbitals compared to the case of Cl, and therefore, prohibitively high cost of the CVS-ADC(2)x computation. In our calculations the core space comprises the 1s orbital of K$^{+}$ or \cli, whereas the remaining occupied orbitals are included in the valence space. For the calculations of the XAS spectra we used the C$_2$ point group in the case of K$^{+}$(H$_2$O)$_6$ and \cli(H$_2$O)$_6$. To account for the lifetime broadening due to the Auger decay of the core excited states, we convolved the theoretical spectra with a Lorentzian function of FWHM 0.74\,eV and 0.62\,eV in the case of \ki~and \cli, respectively \citep{Krause79:329}. Additionally, we convolved the theoretical spectra with a Gaussian profile to also account for the experimental resolution (see Fig.\ \ref{fg:xas_kcl}). We analyzed the core excited states by expanding the natural orbitals occupied by the excited electron (singly occupied natural orbitals, SONOs) $\psi_{i}$ of the microsolvated clusters in the basis of SONOs of the bare K$^{+}$ or \cli~ions, $\chi_{nl}$
%
\begin{equation}\label{eq:sono_proj}
\psi_{i} = \sum_{nl} a^{i}_{nl} \chi_{nl}
\end{equation}
%
where $n$ and $l$ stand for the principal and orbital quantum numbers as described in Ref.\ \citep{miteva16:16671}. The expansion coefficients $a^{i}_{nl}$ show the degree of delocalization of the excited electron and the mixing of the core excited states in the ligand field created by the surrounding water molecules (see Fig.\ \ref{fg:xas_kcl}).

The final states following KLL resonant Auger decay of K$^{+}$(H$_2$O)$_6$ and \cli(H$_2$O)$_6$ were computed at the Configuration Interaction Singles (CIS) level using the Graphical Unitary Group Approach (GUGA) as implemented in the GAMESS-US package \citep{GUGA_PhysScr_21,GUGA_JCP_70,GUS}. In order to account for the relaxation effects upon core ionization, we employed a restricted open-shell Hartree-Fock reference wave function with a hole in the 2s orbital of both \ki~and \cli.  We used the 6-311++G(2d,2p) basis set \citep{Krishnan80:650,McLean80:5639,Blaudeau97:5016} on all atoms. Additionally, the basis set was augmented with two sets of s, p, d diffuse functions in the case of \ki, and three sets of s, p, d diffuse functions in the case of \cli. The larger basis set employed in the case of Cl was necessary in order to ensure the convergence of the excited states. The active space comprises the 2s and 2p orbitals of K/Cl with occupancy fixed to 6 and all virtual orbitals with occupancy fixed to 1. The remaining doubly occupied orbitals were frozen in the calculation. \citep{mosnier16:061401}
